---
title: "A database of normal human tissue methylation values: `methdbr`"
author: "Florent Chuffart & Magali Richard"
date: "`r Sys.Date()`"
output: 
  rmarkdown::html_document:
    toc: true
    toc_float: true
    toc_depth: 3
    number_sections: true
---


# Description

The `methdbr` package provides a dataset of normal human tissue methylation values. Tissues and methylation are described into a 2 distincts data frames.

# Content of the package

Objects available in the package are:

  * `methdbr_d`   A matrix of 187 x 137595 beta values (methylation).
  * `methdbr_e`   A data frame that describes the 187 samples.
  * `methdbr_pf`  A data frame that describes the 137595 CpG probes.
  
# Installation

To get the current development version from github:

```
devtools::install_github("fchuffar/methdbr")
```

# Usage

```
rmarkdown::render("README.Rmd")
```



# Exemple: inputation of missing sex values

```{r, echo=FALSE, eval=TRUE}
knitr::opts_chunk$set(collapse=TRUE, comment = "#>", fig.width=9, fig.height=6, eval=TRUE, echo=FALSE, results="hide")
```

## Loading data
```{r echo=TRUE}
d  = methdbr::methdbr_d 
e  = methdbr::methdbr_e 
pf = methdbr::methdbr_pf
```


## Sex status

```{r echo=TRUE, results="verbatim"}
e$sex
```

## Logistic Regression

```{r echo=TRUE}
s = as.numeric(as.factor(e[colnames(d),]$sex)) - 1
pvals = apply(d[8000:8500,], 1, function(meth){
  # meth = d["cg22507154",]
  m = glm(s~meth, family = binomial(logit))
  m$coefficients
  summary(m)$coefficients
  pval = summary(m)$coefficients[2,4]
})

plot(-log10(pvals), main="Manhattan plot")

probes = names(pvals)[order(pvals)][1:6]


layout(matrix(1:6, 2, byrow=TRUE), respect=TRUE)
for (probe in probes[1:6]) {
  meth = d[probe,]
  plot(meth, s, main=paste0("s~meth ", probe), col=s+1)
  m = glm(s~meth, family = binomial(logit))
  logitinv = function(x) 1/(1 + exp(-x))
  x = sort(meth)
  lines(x, logitinv(m$coefficients[[1]] + m$coefficients[[2]]*x), col=2, lwd=2)
  py1x = function(t,m) {
    x = m$coefficients[[1]] + m$coefficients[[2]]*t
    1/(1 + exp(-x))
  }
  suppressWarnings(arrows(meth, s, meth, py1x(meth,m), col=adjustcolor(4, alpha.f=0.2), length=0.05, lwd=3))
  legend("bottomright","P(Y|X)", col=4, lty=1, cex=0.6)  
}

```

## Imputation of missing sex values

```{r echo=TRUE}
# TODO
```

# Session Information

```{r, results="verbatim"}
sessionInfo()
```

